Four-body L=1 systems from contact EFT

### 2022-03-27 manuscript/Draft_0.3.tex
India V1

\usepackage{epsfig}
\usepackage[utf8]{inputenc}
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
%\usepackage{slashed}
\usepackage{caption}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{bbold}
\usepackage{amssymb,latexsym}
\usepackage{amsmath,amsbsy,bbm}

@@

\newcommand{\eftnopi}{\mbox{EFT$(\not \! \pi)$}}

@@

\section{Appendix: Resonating group method and interactions}
In this section, we reformulate the $A$-body problem of one particle scattering elastically off
an $(A-1)$-body bound state with a relative kinetic energy which is small relative to the lowest
excitation energy of the target compound.

>>>>

\usepackage{graphicx} 
\usepackage{dashrule}
\usepackage{dcolumn} 
\usepackage{longtable} 
\usepackage{mathtools}
\usepackage{amssymb} 
\usepackage{bbold}
\usepackage{amsmath} 
\usepackage{amsfonts} 
\usepackage{wasysym}
\usepackage{slashed} 
\usepackage[dvipsnames]{xcolor} 
\usepackage{soul}
\usepackage{rotating}
\usepackage{paracol}
\usepackage{nicefrac}
\usepackage{MnSymbol}

@@

\newcommand{\eg}{\textit{e.g.}~}
\newcommand{\ie}{\textit{i.e.}~}
\newcommand{\viz}{\textit{viz.}~}
\newcommand{\cf}{\textit{cf.}~}
\newcommand{\eftnopi}{\mbox{EFT$(\not \! \pi)$}}
\newcommand{\ve}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\rcm}{\ensuremath{\ve{R}_\text{\scriptsize{c.m.}}}}
\newcommand*{\mprime}{^{\prime}\mkern-1.2mu}
\newcommand*{\mdprime}{^{\prime\prime}\mkern-1.2mu}
\newcommand*{\mtprime}{^{\prime\prime\prime}\mkern-1.2mu}
\newcommand{\ddrei}[1]{\delta_{\tiny \lambda}^{(3)}\!\big(#1\big)}
\newcommand{\cc}{\ensuremath C(\lambda)}
\newcommand{\dd}{\ensuremath D(\lambda)}

\newcommand{\bra}[1] {\left\langle~#1~\right|}
\newcommand{\ket}[1] {\left|~#1~\right\rangle}
\newcommand{\bet}[1] {\left|#1\right|}
\newcommand{\overlap}[2] {\left\langle\,#1\,\left|\,#2\,\right.\right\rangle}
\newcommand{\me}[3] {\left\langle\,#1\,\left.\left|\,#2\,\right|\right.\,#3\,\right\rangle}
\newcommand{\mer}[3] {\left\langle\,#1\,\left|\left.\,#2\,\right|\,#3\,\right.\right\rangle_{\ve{R}}}
\newcommand{\redme}[3] {\left\langle\,#1\,\middle|\right|\,#2\,\left|\middle|\,#3\,\right\rangle}

@@

\section{Appendix: Resonating group method and interactions}

%We treat $A$ as the triton, $B$ as the neutron, and expand the cluster-internal function in a Gaussian basis
%
%\begin{equation}\label{eq.app.trit}
%\phi_3=\sum_n^{N_f}\,c_n\cdot e^{-\alpha_n\sum_{i=1}^{3}\left(\ve{r}_i-\ve{R}_3\right)^2}
%\qquad
%\begin{array}{ll}
%     N_f      & \text{\small : basis dimension}  \\
%     \ve{r}_i & \text{\small : single-particle coordinates}  \\
%     \ve{R}_3 & \text{\small : center of mass of the triton fragment}
%\end{array} 
%\quad,
%\end{equation}
%
%to allow for an analytical reduction of the equation of motion from its full four-body to an
%effective two-body version which assumes that the triton-internal motion is unaffected by the
%scattered neutron.
%
% As we treat the triton as inert, we consider

The non-relativistic dynamics of the four particles is given by the solutions to

\begin{equation}\label{eq.rgm.eqom}
\left(-\frac{\hbar^2}{2\mu}\ve{\Delta}_R-E+\mathcal{V}\right)\ket{\Psi}
~
\Leftrightarrow
~
\me{\delta\Psi}{\hat{H}-E}{\Psi}
=
0
\qquad.
\end{equation}

The second, viz. the projection/variational equation is equivalent to the Schr\"odinger equation if
$\delta\Psi$ denotes an arbitrary variation of the state.
The resonating-group methods expresses the total wave function of this state in the form

\begin{eqnarray}\label{eq.app.rgmwfkt}
\Psi
&=&
\mathcal{A}\Big\lbrace
\sum_i\phi(A_i)\,\phi(B_i)\,\chi_i(\ve{R}_i)
\nonumber\\
&&
+\sum_j\phi(A_j)\,\phi(B_j)\,\phi(C_j)\,\chi_j(\ve{R}_{j1},\ve{R}_{j2})
+\sum_k\phi(A_k)\,\phi(B_k)\,\phi(C_k)\,\phi(D_k)\,\chi_j(\ve{R}_{k1},\ve{R}_{k2},\ve{R}_{k3})
\nonumber\\
&&
+\ldots+\sum_mc_m\eta_m
\Big\rbrace~
\zeta(\rcm)
\qquad,
\end{eqnarray}

where each term corresponds to a particular fragmentation of the particles into two, three, four, $\ldots$ cluster.
The internal motion of such a cluster of $A$ particles is expanded with a complete set of functions $\phi(A_i)$, and the
relative motion between the clusters is encoded in the $\chi_i$'s. The last term is included to improve the expansion if the
fragment-internal basis is not complete and/or if the cluster expansion is truncated such that a more complicated shape
of the state for small separations between the clusters cannot be accounted for.

For the elastic scattering of a neutron off a three-nucleon targets whose ground state is 
$\Delta\epsilon_\text{nuclear}\approx6.2~$MeV and
$\Delta\epsilon_\text{unitary}\approx8.4~$MeV, respectively, below the lowest dissociation thresholds,
and the relative energies between the fragments $E_r\ll\Delta\epsilon$,
we chose to express the wave function of the system with a two-cluster, single-channel ansatz
without specific distortion:

\begin{equation}\label{eq.app.rgmwfkt2}
\Psi
=
\mathcal{A}_{31}\left[\phi(3)\phi(1)\,\chi(\ve{R})\,\zeta(\rcm)\right]
=
\int\mathcal{A}_{31}\left[\phi(3)\phi(1)\,\delta\big(\ve{R}-\ve{R}\mprime\big)\,\zeta(\rcm)\right]\,\chi(\ve{R}\mprime)\,d\ve{R}\mprime
\qquad.
\end{equation}

\cf\eqref{eq.app.rgmwfkt}, first term. The parameter representation employed with the second equality in \eqref{eq.app.rgmwfkt2}
will yield a defining equation for the relative motion independent of the antisymmetrizer $\mathcal{A}$.
Assuming an inert triton core allows solely for a variation of the relative wave function:

\begin{equation}\label{eq.app.rgmwfkt}
\delta\Psi
=
\int\,\delta\chi(\ve{R}\mprime)\,\mathcal{A}_{AB}\left[\phi_A\phi_B\,\delta\big(\ve{R}-\ve{R}\mprime\big)\,\zeta(\rcm)\right]\,d\ve{R}\mprime
\equiv
\sumint \delta a_n\Phi_n
\qquad.
\end{equation}

The latter notation allows for the following concise form of the projection equation

\begin{equation}\label{eq.app.rgmwfkt}
\me{\sumint \delta a_m\Phi_m}{\hat{H}-E}{ \sumint a_n\Phi_n}=0
~\to~\mer{\phi(3)}{\hat{H}-E}{\mathcal{A}_{31}\left[\chi(\ve{R}\mprime)\,\phi(3)\right]}=0
\qquad
\end{equation}

where the {\it ket} subscript average indicates that the average over the relative radius was already taken,
and the remaining integrals are over the fragment-internal coordinates. If the cluster-internal states are properly
anti-symmetrized, it remains to consider the inter-fragment anti-symmetrization via
 $\mathcal{A}_{31}=\mathbb{1}-\hat{P}_{14}$.
We chose nucleons 1 and 4 to be the neutrons with identical spin orientation, and hence need to consider the permutation of
those particles, only. All other permutations $\hat{P}_{ij}$ will yield zero matrix elements because of the orthogonality of
the internal states.

Being more explicit, the motion of the two fragments relative to each other is approximated by
\begin{equation}\label{eq.rgm.eqom}
\int\left\lbrace~\phi(3)\left(-\frac{\hbar^2}{2\mu}\ve{\Delta}_R-E+B(3)+\mathcal{V}_{31}\right)
\mathcal{A}_{31}\left[\phi(3)\,\chi(\ve{R})\right]\right\rbrace d\ve{r}_{1,2,3}=0
\qquad.
\end{equation}

Employing the parameter representation, $\chi(\ve{R})=\int\delta(\ve{R}-\ve{R}\mprime)\chi{\ve{R}\mprime}d\ve{R}\mprime$,
and expanding the triton ground state in Gaussians,
$\phi(3)=\sum_n^{N_f}\,c_n\cdot e^{-\alpha_n\sum_{i=1}^{N_A}\left(\ve{r}_i-\ve{R}_3\right)^2}$,
in combination with the leading-order \eftnopi~potential

\begin{equation}\label{eq.app.pot}
\mathcal{V}_{31}=\cc\,\sum_{i\in A\atop j\in B}\,\ddrei{\ve{r}_i-\ve{r}_j}+
\dd\,\sum_{i,j,k\atop i\in A\Rightarrow j\vee k\in B}\ddrei{\ve{r}_i-\ve{r}_j}\,\ddrei{\ve{r}_i-\ve{r}_k}
\end{equation}

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{./Graphs/rgm_potential} 
\caption{Two-fragment potential as a function of the distance between a
neutron and the center of mass of a triton. The plot shows the $P$-wave component of the
interaction including local and non-local part of the kinetic-, pair-, and three-body interactions.
The non-local parts are evaluated at $\rho=\rho'$}
\label{fig:rgm_potential}
\end{figure}
